---
title: "P8105 Homework 1"
author: Aisha Waggeh
date: September 20 2025
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(moderndive)
library(ggplot2)
library(tidyverse)
```


```{r load-data}
# Load the early_january_weather dataset
data("early_january_weather")
```

I'm an R Markdown document! 

# Problem 0.1

  This .Rmd file is part of my GitHub repo for Homework 1.
  It renders to github_document, and the repo is organized as instructed.

# Problem 0.2

I will follow good coding style by:

*   using clear variable names,
*   writing one command per line,
*   adding white space for readability,
*   and explaining results in plain text.

# Problem 1: Exploring Early January Weather
We will first explore the 'early_january_weather' dataset.
I rename data set to weather_df

```{r}
data("early_january_weather")
weather_df = early_january_weather
```

*    The dataset has `r nrow(weather_df)` rows and `r ncol(weather_df)` columns.  
*    The variables are: `r paste(names(weather_df), collapse = ", ")`.  
*    The mean temperature is `r round(mean(weather_df$temp, na.rm = TRUE), 2)` °F.

#I made a scatterplot of temp (y) vs time_hour (x); color points using humid variable & adding color inside of aes in my ggplot.

```{r}
ggplot(weather_df, aes(x = time_hour, y = temp, color = humid)) +
  geom_point(alpha = 0.7) +
  labs(
    title = "Temperature vs Time (colored by Humidity)",
    x = "Time (hourly)",
    y = "Temperature (F)",
    color = "Humidity"
  )

#  Display and save the plot to my project directory

 ggsave("temp_vs_humid.png", width = 6, height = 4)
```

*Observations from the scatter plot*

*    Temperature fluctuates across the observed hours between Jan 07 and Jan 14.
*    Points are colored by humidity. Higher humidity tends to occur at higher temperatures.
*    Overall, the scatterplot shows how temperature and humidity vary together during early January. 

# Problem 2
*In this problem, I created a data frame with four types of variables:

1.  A random sample of size 10 from a standard Normal distribution.
2.  A logical vector indicating whether each sample value is greater than 0.
3.  A character vector of length 10.
4.  A factor vector of length 10, with three different factor levels.


```{r}
set.seed(123) # this numbers for reproducibility
  
#create variables
  sample_vec = rnorm(10)
  logical_vec = sample_vec > 0
  char_vec = c("a", "b", "c", "d", "e", "f", "g","h", "i", "j")
  factor_vec = factor(rep(c("low", "medium", "high"), length.out = 10))
  
  
#combine into a dataframe
    df = data.frame(
      sample = sample_vec,
      logical= logical_vec,
      character = char_vec,
      factor = factor_vec
    )
    df
```

#Now I will try to take the mean of each variable in my dataframe. 

```{r}
library(tidyverse)

mean(df %>% pull(sample))     # works
mean(df %>% pull(logical))    # works (TRUE = 1, FALSE = 0)
mean(df %>% pull(character))  # will not work b/c non-numeric
mean(df %>% pull(factor))     # will not work b/c non-numeric
```


#Now I will apply 'as.numeric' to the logical, character, and factor variables created.

```{r}
as.numeric(df$logical)   # TRUE becomes 1, FALSE becomes 0
as.numeric(df$character) # NA with a warning, because characters cannot be converted to numbers
as.numeric(df$factor)    # returns the internal integer codes used to represent factor levels (not labels themselves)
```

#Now on to what is happening in my code chunk when I apply the 'as.numeric' to the above:

*   Logical vector when we coerce it with 'as.numeric' and take the mean of the logical colunm works, because     it gives the proportion of values that are TRUE (the proportion in my sample greater than 0).

*   Character vector such as "a" "b" "c" to numeric produces NA with a warning, because there is no sensible     way to convert letters to numbers. 

*   Factor vector when we coerce it with 'as.numeric' returns its integer codes such as 1 for "low", 2 for       "medium" and 3 for "high".

*   Lastly, taking the mean of these code would just be averaging aribtary integers (which has nothing to do     with the factors labels). This explains why the mean(factor) did not work directly.  
