---
title: "Simple document"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(moderndive)
library(ggplot2)
library(tidyverse)
```


```{r load-data}
# Load the early_january_weather dataset
data("early_january_weather")
```

I'm an R Markdown document! 

Problem 0.1

  This .Rmd file is part of my GitHub repo for Homework 1.
  It renders to github_document, and the repo is organized as instructed.

Problem 0.2

I will follow good coding style by:
  using clear variable names,
  writing one command per line,
  adding whitespace for readability,
  and explaining results in plain text.

# Problem 1: Exploring Early January Weather
We will first explore the 'early_january_weather' dataset.


```{r}
data("early_january_weather")
weather_df = early_january_weather
```

*    The dataset has `r nrow(weather_df)` rows and `r ncol(weather_df)` columns.  
*    The variables are: `r paste(names(weather_df), collapse = ", ")`.  
*    The mean temperature is `r round(mean(weather_df$temp, na.rm = TRUE), 2)` °F.

#I made a scatterplot of temp (y) vs time_hour (x); color points using humid variable & adding color inside of aes in my ggplot.

```{r}
ggplot(weather_df, aes(x = time_hour, y = temp, color = humid)) +
  geom_point(alpha = 0.7) +
  labs(
    title = "Temperature vs Time (colored by Humidity)",
    x = "Time (hourly)",
    y = "Temperature (F)",
    color = "Humidity"
  )

#  Display and save the plot to my project directory

 ggsave("temp_vs_humid.png", width = 6, height = 4)
```

# Problem 2

I can take the mean of the sample, too!